Fused regression for network inference in multiple species

### 2015-05-14 paper1.tex
hi

\subsection{Model selection}

@@

\section{Bacterial data}
We used gene-expression data from B. Subtilis and B. Anthracis in order to assess performance gains of fused regression on real data. Our B Subtilis data set consists of 266 time-series and steady-state observations of 4891 genes during development. Our B. Anthracis dataset consists of 72 time-series and steady-state observations of 5536 genes comprising data from developmental and iron-starvation conditions. There were 247 known Transcription Factors (TFs) in the B. Subtilis dataset, and 248 TFs in the B. Anthracis dataset.

We obtained 1870 one-to-one orthologs from Inparanoid, which produced 177650 fusion-constraints between gene interactions within the two species. Although this number seems very large, it represents only $14.7\%$ of the regulatory interaction matrix in B. Subtilis and $12.9\%$ in B. Anthracis. The use of such dissimilar species is a test of whether fusion can improve network-inference even when the overlap in interactions is very small.

In order to evaluate network inference performance, and for use as priors, we used a gold standard of 2896 known B. Subtilis interactions with corresponding activation and repression sign. Of these 2896 priors, 968 had corresponding interactions in B. Anthracis. Based on our simulation results, we can expect the greatest gains in network-inference performance from fusion when the species of interest has a small number of available conditions, but data is abundant in a related species. However, in order to evaluate performance objectively a gold-standard of known interactions is necessary. As a result, we can only evaluate network recovery for B. Subtilis, and B. Subtilis also has the majority of our conditions. In order to simulate the data-poor regime, we subsampled our B. Subtilis data. Specifically, we divided our B. Subtilis data into $k$ folds, and then for each fold fit a network to the B. Subtilis data from that fold alone fused to the entire 72 B. Anthracis conditions. We can then compute performance metrics (ie AUPR) for each fold, and use their variability across folds to test for significance of any improvements relating to the use of fusion constraints.

\subsection{Bacterial performance}
\subsubsection{Performance with priors}
We want to show that performance increases with introduction of priors - this is easy/circular but probably important.

>>>>

\subsection{Model selection}
We can define a leave-out set of priors and choose parameters based on cross-validated AUPR on the leave-out set.

@@

\section{Bacterial data}
We used gene-expression data from B. subtilis and B. anthracis in order to assess performance gains of fused regression on real data. Our B subtilis data set consists of 266 time-series and steady-state observations of 4891 genes during development. Our B. anthracis dataset consists of 72 time-series and steady-state observations of 5536 genes comprising data from developmental and iron-starvation conditions. There were 247 known Transcription Factors (TFs) in the B. subtilis dataset, and 248 TFs in the B. anthracis dataset.

We obtained 1870 one-to-one orthologs from Inparanoid, which produced 177650 fusion-constraints between gene interactions within the two species. Although this number seems very large, it represents only $14.7\%$ of the regulatory interaction matrix in B. subtilis and $12.9\%$ in B. anthracis. The use of such dissimilar species is a test of whether fusion can improve network-inference even when the overlap in interactions is very small.

In order to evaluate network inference performance, and for use as priors, we used a gold standard of 2896 known B. subtilis interactions with corresponding activation and repression sign. Of these 2896 priors, 968 had corresponding interactions in B. anthracis. Based on our simulation results, we can expect the greatest gains in network-inference performance from fusion when the species of interest has a small number of available conditions, but data is abundant in a related species. However, in order to evaluate performance objectively a gold-standard of known interactions is necessary. As a result, we can only evaluate network recovery for B. subtilis, and B. subtilis also has the majority of our conditions. In order to simulate the data-poor regime, we subsampled our B. subtilis data. Specifically, we divided our B. subtilis data into $k$ folds, and then for each fold fit a network to the B. subtilis data from that fold alone fused to the entire 72 B. anthracis conditions. We can then compute performance metrics (ie AUPR) for each fold, and use their variability across folds to test for significance of any improvements relating to the use of fusion constraints.

\subsection{Bacterial performance}
\subsubsection{Performance with priors}
We want to show that performance increases with introduction of priors - this is easy/circular but important to show that our method can effectively incorporate priors

\subsubsection{Adaptive fusion}
We want to show that we can successfully identify neofunctionalization; since we do not have a gold standard for this we will introduce false positive orthologs into the orthology information and evaluate the ability of our method to identify true from false orthologs.
